---
title: "EDA for Wine Quality Dataset"
author: "Ammar Alfaifi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
# knitr::knit(function(input, ...) {
#     rmarkdown::render(
#       input,
#       output_dir = "../"
#     )
#   })
```

# Introduction

This a report about the wine quality dataset downloaded form kaggle api.
In this report we will examine this dataset and explore it to find out its significant
as well as the research questions that might answer. This approach is commonly called 
Exploratory Data Analysis (in short, EDA).

# Data
As described form the kaggle sourse in [source](https://www.kaggle.com/datasets/uciml/red-wine-quality-cortez-et-al-2009), 
this datasets is related to red variants of the Portuguese "Vinho Verde" wine.

## Describing the headers

| fixed acidity        | most acids involved with wine or fixed or nonvolatile (do not evaporate readily)                                              |
|----------------------|-------------------------------------------------------------------------------------------------------------------------------|
| volatile acidity     | the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste                       |
| citric acid          | found in small quantities, citric acid can add 'freshness' and flavor to wines                                                |
| residual sugar       | the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and               |
| chlorides            | the amount of salt in the wine                                                                                                |
| free sulfur dioxide  | the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents          |
| total sulfur dioxide | amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2             |
| density              | the density of water is close to that of water depending on the percent alcohol and sugar content                             |
| pH                   | describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the |
| sulphates            | a wine additive which can contribute to sulfur dioxide gas (S02) levels,                                                      |

# EDA

```{r, include=FALSE}
library(tidyverse)
library(Hmisc)
library(funModeling) 
library(DataExplorer)
library(janitor)
library(here)
library(DT)
```

Loading and cleaning the header names
```{R}
read_csv(here("data/winequality-red.csv")) %>%
    clean_names() -> wine_data
glimpse(wine_data)
```

```{r, }
status(wine_data)
```

This shows a summury about the dataset, includin standard deviation, mean, and ration on `NA`
values out of total observations.

```{r, }
plot_num(wine_data)
```

The above graphs show the count of observations for each catogerical along with their values.
We can note that for `chlorides` thre is string positve skweness. This might indicat 
unnormal value. Further for `total_sulfur_dioxide`, it has relatively large sdandard deviation.
This can be due to imprecise measuring. Let us investigate more about the later.

```{r}
profiling_num(wine_data$total_sulfur_dioxide)
```

The variation_coef is about 0.7, wich indicate hight uncertinity about whether the 
total_sulfur_dioxide has that effect on the wine quality. It us inresting question to 
analize more about this connection.

## Show Some Catogericals

```{r}
wine_data %>%
    select(chlorides, alcohol, quality) -> selected_table
datatable(selected_table)
```

# Conclusion

to sum up, this dataset has many catogericals that are intresting to analize and study.Furthermore, it 
can leads to a good understanding of how different compoents can increase the wine quality.
More analysis hypothsis testing is needed to determine best practice for wine manufacturers
to produce high wine quality with balance.